HMMotion: Using tracking data to predict coverage strategies in American Football

### 2024-12-31 submission.Rmd
sub R2

## Data
We analyze tracking data from nine weeks of the NFL 2022 season, provided by the NFL Big Data Bowl 2024. Beside the tracking data, we also use information on plays and players. We further considered the corresponding data from PFF that assigned the categories \textit{man}, \textit{zone} and \textit{others} representing the different schemes to each play. As it is not properly described what \textit{others} means, we omit every play that is associated with this value. Moreover, we omit plays with more than five offensive linemen and with two quarterbacks and those plays that did not contain any pre-snap motion. Then, we end up with XY offensive plays in total, from which the defense played Y in zone and X in man coverage.

##### Feature engineering
To accurately forecast the defensive scheme (man- or zone defense) for every play, we need to create various features derived from the tracking data. In particular, we conducted the following feature engineering steps: We first consider all 11 players on each side of the field and compute features related to the convex hull of the positions of the players. In particular, for defense and offense, we compute the area spanned by the convex hull of all player such as well as the largest $y$ distance (i.e.\ the width of the hull) and the largest $x$ distance (i.e. the length of the hull). In addition, we select the five most relevant players on each side of the field. For offense, we omit the offensive line and the QB, while, for defense, we omit nose tackles, defensive tackles and defensive ends, and select the five defenders that were the closest to the five attackers corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis. Finally, we use their standardized $x$ and $y$ coordinates as covariates and order defensive and offensive players according to their $y$ coordinates.. (`r colorize("Distances, orientation?? Appenix?? Rouven: würde sagen ja","red")`)

@@

#### 1. Pre-motion prediction
We train different models (LASSO, Random Forest, XGBoost) to predict whether the defense plays a man- or zone coverage scheme. In particular, .....

>>>>

## Data
We analyze tracking data from nine weeks of the NFL 2022 season, provided by the NFL Big Data Bowl 2024. Beside the tracking data, we also use information on plays and players. We further considered the corresponding data from PFF that assigned the categories \textit{man}, \textit{zone} and \textit{others} representing the different schemes to each play. As it is not properly described what \textit{others} means, we omit every play that is associated with this value. Moreover, we omit plays with more than five offensive linemen and with two quarterbacks `r colorize("and those plays that did not contain any pre-snap motion","orange")` (`r colorize("Ich würde deutlicher drauf eingehen, dass wir nur an motion plays interessiert sind! Vl ein Punkt machen und schreiben: Since we are specifically interested in analyzing pre-snap player movements, we omit plays that did not contain any pre-snap motion.","red")`). Then, we end up with 3985 offensive plays in total, from which the defense played 2973 in zone and 1012 in man coverage.

##### Feature engineering
To accurately forecast the defensive scheme (man- or zone defense) for every play, we need to create various features derived from the tracking data. In particular, we conducted the following feature engineering steps: We first consider all 11 players on each side of the field and compute features related to the convex hull of the positions of the players. In particular, for defense and offense, we compute the area spanned by the convex hull of all player such as well as the largest $y$ distance (i.e.\ the width of the hull) and the largest $x$ distance (i.e. the length of the hull). In addition, we select the five most relevant players on each side of the field. For offense, we omit the offensive line and the QB, while, for defense, we omit nose tackles, defensive tackles and defensive ends, and select the five defenders that were the closest to the five attackers corresponding to a weighted euclidean distance, putting much more emphasis on the y-axis. Finally, we use their standardized $x$ and $y$ coordinates as covariates and order defensive and offensive players according to their $y$ coordinates. Additionally, for each of the relevant defenders, we compute distances to the football and their orientation with respect to the quarterback (`r colorize("values taken at event lineset","orange")`). Finally, we extract relevant information from play-by-play data, such as quarter, down, yards to go, home and away score and the game clock.

Detailed information can be found in the Appendix.

@@

#### 1. Pre-motion prediction
We train different models to predict whether the defense plays a man- or zone coverage scheme. In particular, in the first model, we only use features available pre-motion.
